Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-16 write-ups/PLDI/paper.tex
Cuts.
Move table.
Bang

The $\offerkill$ primitive allows the cooperative kernel to return
compute units to the scheduler by offering to sacrifice workgroups.
The idea is as follows: allowing the scheduler to arbitrarily and abruptly terminate execution
of workgroups might be drastic, yet the kernel
may contain specific program points at which a workgroup could
\emph{gracefully} leave the computation.  We give examples of these
for work stealing and graph traversal in \mysec\ref{sec:programmingguidelines}.

@@

Suppose a workgroup with id $m$ executes $\offerkill$.  
If the workgroup has the highest largest id of all active workgroups then it can be killed by the scheduler, except that workgroup 0 can never be killed (to ensure that the kernel computation does not terminate early).  More formally, if $m < M-1$
or $M=1$ then $\offerkill$ is a no-op.  If instead $M > 1$ and $m =
M-1$, the scheduler can choose to ignore the offer, so that $\offerkill$
executes as a no-op, or accept the offer, so that execution of the workgroup ceases and
the number of active workgroups $M$ is atomically decremented by one.  In the second case, the scheduler can assign some other task to
the compute unit on which the workgroup was executing.

@@

\begin{table}[h]
\footnotesize
\centering
\begin{tabular}{ l l l l }
Chip & HD520 & HD5500 & Iris\\
\hline
\# CU  & 24 & 24 & 47 \\
Driver   & 20.19.15.4501 & 10.18.15.4281 & 20.19.15.4463 \\
Host CPU & i3-6100U & i7-5600U & i3-5157U \\
\end{tabular}
\caption{The Intel GPU platforms used in our evaluation}
\label{tab:chipstested}
\end{table}

@@

Co-scheduling non-cooperative kernels that compute their workload using a single workgroup leads to almost no overhead, but the period is far too high to meet the needs
of any of our three workloads. For example, a heavy workload averages a
period of 939 ms. As more workgroups are dedicated to non-cooperative
kernels, they execute quickly enough to be scheduled at the expected
period. More concretely, for the light and medium workloads, a quarter
of the workgroups executing the non-cooperative kernel are able to meet
their goal period (70 and 40 ms resp.). However, a quarter of the
workgroups are not sufficient to meet the goal for the heavy workload
(giving a mean period of 88 ms). Finally, if half of the workgroups are
allocated to the non-cooperative kernel, the heavy workload averages \~10\% over its goal period (mean of 44 ms). If asking for all-but-one, the heavy workload can reach its goal period.
%
Yet, as expected, allocating more non-cooperative workgroups increases
the execution time overhead of the cooperative kernel.

>>>>

The $\offerkill$ primitive allows the cooperative kernel to return
compute units to the scheduler by offering to sacrifice workgroups.
The idea is as follows: allowing the scheduler to arbitrarily and abruptly terminate execution
of workgroups might be drastic, yet the kernel
may contain specific program points at which a workgroup could
\emph{gracefully} leave the computation.
%We give examples of these
%for work stealing and graph traversal in \mysec\ref{sec:programmingguidelines}.

@@

Suppose a workgroup with id $m$ executes $\offerkill$.  
If the workgroup has the highest largest id of all active workgroups then it can be killed by the scheduler, except that workgroup 0 can never be killed (to ensure that the kernel computation does not terminate early).  More formally, if $m < M-1$
or $M=1$ then $\offerkill$ is a no-op.  If instead $M > 1$ and $m =
M-1$, the scheduler can choose to ignore the offer, so that $\offerkill$
executes as a no-op, or accept the offer, so that execution of the workgroup ceases and
the number of active workgroups $M$ is atomically decremented by one.  
%In the second case, the scheduler can assign some other task to
%the compute unit on which the workgroup was executing.

@@

\begin{table}[t]
\footnotesize
\centering
\begin{tabular}{ l l l l }
Chip & HD520 & HD5500 & Iris\\
\hline
\# CU  & 24 & 24 & 47 \\
Driver   & 20.19.15.4501 & 10.18.15.4281 & 20.19.15.4463 \\
Host CPU & i3-6100U & i7-5600U & i3-5157U \\
\end{tabular}
\caption{The Intel GPU platforms used in our evaluation}
\label{tab:chipstested}
\end{table}

@@

Co-scheduling non-cooperative kernels that request a single workgroup leads to almost no overhead, but the period is far too high to meet the needs
of any of our three workloads; e.g.\ a heavy workload averages a
period of 939 ms. As more workgroups are dedicated to non-cooperative
kernels, they execute quickly enough to be scheduled at the expected
period. For the light and medium workloads, a quarter
of the workgroups executing the non-cooperative kernel are able to meet
their goal period (70 and 40 ms resp.). However, a quarter of the
workgroups are not sufficient to meet the goal for the heavy workload
(giving a mean period of 88 ms). If half of the workgroups are
allocated to the non-cooperative kernel, the heavy workload averages $\sim$10\% over its goal period (mean of 44 ms). If all-but-one are allocated, the heavy workload reaches its goal period.
%
Yet, as expected, allocating more non-cooperative workgroups increases
the execution time of the cooperative kernel.
